---
title: "Wine Reviews"
author: "Vadym Musiienko"
subtitle: "Analyzing Trends in Taste, Price, and Quality"
date: "29 Nov 2024"
format: 
  revealjs:
    logo: media/logo.png
    footer: "Wine Reviews: Analyzing Trends in Taste, Price, and Quality"
    transition: slide
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(ggrepel)

# Load the data
wine_data <- read_csv("data/winemag-data_first150k.csv")

# Clean up the data
wine_data <- wine_data |> filter(!(is.na(price) | is.na(country))) |>
  mutate(country = as.factor(country)) |>
  select(-c(region_1, region_2, winery))
```

## Data frame information


Columns
```{r}
names(wine_data)

```

Rows
```{r}
glimpse(wine_data)
```

## Average price of wine by country


```{r}
#| fig-height: 7
#| fig-width: 12
#| fig-align: center

wine_data |> 
  group_by(country) |>
  summarize(avr_price = mean(price)) |>
  ggplot(aes(fct_reorder(country, avr_price, .desc = TRUE), avr_price,
             fill = country)) +
  geom_col(color = "black", width = 0.8, show.legend = FALSE) +
  labs(
    title = NULL,
    x = NULL,
    y = "Price in USD"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

:::{style="font-size: 22px;"}
<em>
The countries that produce the most expensive wines are the US, France, England, Hungary, and Luxembourg, with average prices ranging from 40 to 50 USD. On the other hand, the cheapest wine producers are Montenegro, Lithuania, Bulgaria, Bosnia, and Ukraine, where wines typically cost between 10 and 15 USD.
</em>
:::

## Code

```{r}
#| eval: false
#| echo: true

wine_data |> 
  group_by(country) |>
  summarize(avr_price = mean(price)) |>
  ggplot(aes(fct_reorder(country, avr_price, .desc = TRUE), avr_price,
             fill = country)) +
  geom_col(color = "black", width = 0.8, show.legend = FALSE) +
  labs(
    title = "Average price of wine by country",
    x = NULL,
    y = "Price in USD"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Correlation between Price and Quality

```{r}
#| fig-height: 6
#| fig-width: 12
#| fig-align: center

wine_data |>
  filter(price < 500) |>
  group_by(price) |>
  summarize(avr_rating = mean(points)) |>
  ggplot(aes(price, avr_rating)) +
  geom_point(alpha = 0.5) +
  geom_smooth(color = "red") +
  labs(
    title = NULL,
    x = "Price (USD)",
    y = "Average rating (Points)"
  ) +
  theme_minimal()
```

:::{style="font-size: 22px;"}
<em>
There is definitely a correlation between price and rating, with more expensive wines tending to have higher ratings. However, the relationship is stronger in the 0 to 100 USD range, where price significantly impacts quality. Beyond 100 USD, wines are fairly similar in quality, with higher prices offering only a slight improvement.
</em>
:::

## Code

```{r}
#| echo: true
#| eval: false

wine_data |>
  filter(price < 500) |>
  group_by(price) |>
  summarize(avr_rating = mean(points)) |>
  ggplot(aes(price, avr_rating)) +
  geom_point(alpha = 0.5) +
  geom_smooth(color = "red") +
  labs(
    title = "Price vs. Quality",
    x = "Price (USD)",
    y = "Average rating (Points)"
  ) +
  theme_minimal()
```

## Frequency of wine adjectives

```{r}
#| fig-height: 6
#| fig-width: 12
#| fig-align: center

# Create a vector with the adjectives
adjectives <- c("Acidic", "Astringent", "Barnyard", "Buttery", "Chewy",
                "Earthy", "Flabby", "Fruity", "Fruit-forward",
                "Herbaceous", "Jammy", "Juicy", "Musty", "Nutty",
                "Oaky", "Opulent", "Perfumed", "Racy", "Spicy",
                "Supple", "Tannic", "Toasty", "Vegetal", "Velvety")

# Create a regex pattern to search for adjectives
pattern <- paste0("(?i)", paste(adjectives, collapse = "|"))

# Create a data frame with wine reviews where those adjectives were used
wine_with_adj <- wine_data |> 
  mutate(
    adjective = factor(tolower(str_extract(description, pattern)))
    ) |>
  filter(!is.na(adjective))

# Plot the frequency of each adjective
wine_with_adj |> group_by(adjective) |>
  summarise(count = n()) |>
  ggplot(aes(count, fct_reorder(adjective, count, .desc = TRUE), 
             fill = adjective)) +
  geom_col(color = "black", width = 0.8, show.legend = FALSE) +
  labs(
    title = NULL,
    y = NULL,
    x = "Number of times mentioned"
  ) +
  theme_minimal()
```

## Code

```{r}
#| echo: true
#| eval: false

# Create a vector with the adjectives
adjectives <- c("Acidic", "Astringent", "Barnyard", "Buttery", "Chewy",
                "Earthy", "Flabby", "Fruity", "Fruit-forward",
                "Herbaceous", "Jammy", "Juicy", "Musty", "Nutty",
                "Oaky", "Opulent", "Perfumed", "Racy", "Spicy",
                "Supple", "Tannic", "Toasty", "Vegetal", "Velvety")

# Create a regex pattern to search for adjectives
pattern <- paste0("(?i)", paste(adjectives, collapse = "|"))

# Create a data frame with wine reviews where those adjectives were used
wine_with_adj <- wine_data |> 
  mutate(
    adjective = factor(tolower(str_extract(description, pattern)))
    ) |>
  filter(!is.na(adjective))

# Plot the frequency of each adjective
wine_with_adj |> group_by(adjective) |>
  summarise(count = n()) |>
  ggplot(aes(count, fct_reorder(adjective, count, .desc = TRUE), 
             fill = adjective)) +
  geom_col(color = "black", width = 0.8, show.legend = FALSE) +
  labs(
    title = "Adjective Frequency in Wine Reviews",
    y = NULL,
    x = "Number of times mentioned"
  ) +
  theme_minimal()
```

## Average Price and Rating of Wine by Flavor Note

```{r}
#| fig-height: 6
#| fig-width: 12
#| fig-align: center

# Plot average price and rating depending on a flavor note
wine_with_adj |> group_by(adjective) |>
  summarise(avr_price = mean(price), avr_points = mean(points)) |>
  ggplot(aes(avr_price, avr_points)) +
  geom_point(size = 3) +
  geom_label_repel(aes(label = adjective),
                   size = 2.5, color = "black", fill = "lightyellow", 
                   label.size = 0.25, label.padding = 0.25, 
                   box.padding = 0.25) +
  labs(
    title = NULL,
    x = "Average Price (USD)",
    y = "Average Points (Points)"
  ) +
  theme_minimal()
```


## Code

```{r}
#| echo: true
#| eval: false

wine_with_adj |> group_by(adjective) |>
  summarise(avr_price = mean(price), avr_points = mean(points)) |>
  ggplot(aes(avr_price, avr_points)) +
  geom_point(size = 3) +
  geom_label_repel(aes(label = adjective),
                   size = 2.5, color = "black", fill = "lightyellow", 
                   label.size = 0.25, label.padding = 0.25, 
                   box.padding = 0.25) +
  labs(
    title = "Average Price and Rating of Wine by Flavor Note",
    x = "Average Price (USD)",
    y = "Average Points (Points)"
  ) +
  theme_minimal()
```

